### 2024-08-14 index.qmd
docs(quarto): add basic/empty quarto site

---
authors:
  - name: Amy Heather
    url: https://github.com/amyheather
    orcid: 0000-0002-6596-3479
    affiliation: University of Exeter
---

Summary of the six computational reproducibility assessments conducted as part of STARS Work Package 1. These assessed six discrete-event simulation papers with models in Python and R.