### 2024-08-14 index.qmd
docs(quarto): reworked into a book with empty pages to complete

---
title: Overview
authors:
  - name: Amy Heather
    url: https://github.com/amyheather
    orcid: 0000-0002-6596-3479
    affiliation: University of Exeter
---

TODO: Overview page