### 2024-08-14 pages/paper_evaluation.qmd

---
title: "Evaluation of the article"
---

TODO: Add evaluation of article against STRESS-DES and against the guidelines derived from ISPOR-SDM. Include record of timings to complete this part of the evaluation for each.