### 2024-08-14 pages/background.qmd
docs(quarto): add empty page for background

---
title: "STARS"
---

TODO: Add background about STARS, pilot work, and aim of WP1