### 2024-08-19 pages/background.qmd
docs(intro): add to introduction and methods a little

---
title: "Introduction"
bibliography: ../references.bib
---

<!-- TODO: Write an introduction -->

Review: @monks_computer_2023

Framework: @monks_towards_2024

![STARS framework overview](../images/stars_framework_overview.png)

Aim of this work: Computational reproducibility assessments of six discrete event simulation models in Python and R.

## References