### 2024-08-19 pages/protocol.qmd

---
title: "Methods"
bibliography: ../references.bib
---

<!-- TODO: Write out a summary here -->

Protocol: @heather_protocol_2024

![Workflow for STARS work package 1](../images/stars_wp1_workflow.png)

## References