### 2024-08-23 framework/stars_checklist.tex
docs(stars): example checklist document with conversions

\hypertarget{stars-framework-checklist}{%
\section{STARS Framework Checklist}\label{stars-framework-checklist}}

\hypertarget{essential-components}{%
\subsection{Essential components}\label{essential-components}}

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.47\columnwidth}\raggedright
Item\strut
\end{minipage} & \begin{minipage}[b]{0.47\columnwidth}\raggedright
Implementation\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Open license}\emph{Free and open-source software (FOSS) license
(e.g.~MIT, GNU Public License (GPL))}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Dependency management}\emph{Specify software libraries, version
numbers and sources (e.g.~dependency management tools like virtualenv,
conda, poetry)}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{FOSS model}\emph{Coded in FOSS language (e.g.~R, Julia,
Python)}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Minimum documentation}\emph{Minimal instructions (e.g.~in
README) that overview (a) what model does, (b) how to install and run
model to obtain results, and (c) how to vary parameters to run new
experiments}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{ORCID}\emph{ORCID for each study author}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Citation information}\emph{Instructions on how to cite the
research artefact (e.g.~CITATION.cff file)}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Remote code repository}\emph{Code available in a remote code
repository (e.g.~GitHub, GitLab, BitBucket)}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.47\columnwidth}\raggedright
\textbf{Open science archive}\emph{Code stored in an open science
archive with FORCE11 compliant citation and guaranteed persistance of
digital artefacts (e.g.~Figshare, Zenodo, the Open Science Framework
(OSF), and the Computational Modeling in the Social and Ecological
Sciences Network (CoMSES Net))}\strut
\end{minipage} & \begin{minipage}[t]{0.47\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{optional-components}{%
\subsection{Optional components}\label{optional-components}}

\begin{longtable}[]{@{}lll@{}}
\toprule
\begin{minipage}[b]{0.30\columnwidth}\raggedright
Item\strut
\end{minipage} & \begin{minipage}[b]{0.30\columnwidth}\raggedright
Description\strut
\end{minipage} & \begin{minipage}[b]{0.30\columnwidth}\raggedright
Implementation\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Enhanced documentation\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Open and high quality documentation on how the model is implemented and
works (e.g.~via notebooks and markdown files, brought together using
software like Quarto and Jupyter Book). Suggested content includes:•
Plain english summary of project and model• Clarifying license• Citation
instructions• Contribution instructions• Model installation
instructions• Structured code walk through of model• Documentation of
modelling cycle using TRACE• Annotated simulation reporting guidelines•
Clear description of model validation including its intended
purpose\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Documentation hosting\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Host documentation (e.g.~with GitHub pages, GitLab pages, BitBucket
Cloud, Quarto Pub)\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Online coding environment\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Provide an online environment where users can run and change code
(e.g.~BinderHub, Google Colaboratory, Deepnote)\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Model interface\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Provide web application interface to the model so it is accessible to
less technical simulation users\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.30\columnwidth}\raggedright
Web app hosting\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
Host web app online (e.g.~Streamlit Community Cloud, ShinyApps
hosting)\strut
\end{minipage} & \begin{minipage}[t]{0.30\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}