### 2024-09-03 pages/reproduction.qmd
docs(reproduction): make study descriptions more consistent (e.g. mentioning model language and packages each time)

---
title: "Reproduction"
echo: False
bibliography: ../references.bib
---

## Studies

**@shoaib_simulation_2021:** Uses **python (salabim)** to model primary health centres (PHCs) in India. The model has four patient types: outpatients, inpatients, childbirth cases and antenatal care patients. Four model configurations are developed based on observed PHC practices or government-mandated operational guidelines. The paper explores different operational patterns for scenarios where very high utilisation was observed, to explore what might help reduce utilisation of these resources. Note: The article was as @shoaib_simulation_2022, but we used the green open access pre-print @shoaib_simulation_2021. <a href="https://pythonhealthdatascience.github.io/stars-reproduce-shoaib-2022/evaluation/reproduction_success.html" target="_blank">Link to reproduction.</a>

**@huang_optimizing_2019:** Uses **R (simmer)** to model an endovascular clot retrieval (ECR) service. ECR is a treatment for acute ischaemic stroke. The model includes the stroke pathway, as well as three other pathways that share resources with the stroke pathway: an elective non-stroke interventional neuroradiology pathway, an emergency interventional radiology pathway, and an elective interventional radiology pathway. The paper explores waiting times and resource utilisation - particularly focussing on the biplane angiographic suite (angioINR). A few scenarios are tried to help examine why the wait times are so high for the angioINR. <a href="https://pythonhealthdatascience.github.io/stars-reproduce-huang-2019/evaluation/reproduction_success.html" target="_blank">Link to reproduction.</a>

**@lim_staff_2020:** Uses **python (numpy and pandas)** to model the transmission of COVID-19 in a laboratory. It examines the proportion of staff infected in scenarios varying the: number of shifts per day; number of staff per shift; overall staff pool; shift patterns; secondary attack rate of the virus; introduction of protective measures (social distancing and personal protective equipment). <a href="https://pythonhealthdatascience.github.io/stars-reproduce-lim-2020/evaluation/reproduction_success.html" target="_blank">Link to reproduction.</a>

**@kim_modelling_2021:** Adapts a previously developed **R (Rcpp, expm, msm, foreach, iterators, doParallel)** model for abdominal aortic aneurysm (AAA) screening of men in England. The model is adapted/used to explore different approaches to resuming screening and surgical repair for AAA, as these survives were paused or substantially reduced during COVID-19 due to concerns about virus transmission. <a href="https://pythonhealthdatascience.github.io/stars-reproduce-kim-2021/evaluation/reproduction_success.html" target="_blank">Link to reproduction.</a>

**@anagnostou_facs-charm_2022:** This paper includes two models - we have focussed just on the dynamiC Hospital wARd Management (CHARM) model. CHARM uses **Python (simpy)** to model intensive care units (ICU) in the COVID-19 pandemic (as well as subsequent stays in a recovery bed). It includes three types of admission to the ICU (emergency, elective or COVID-19). COVID-19 patients are kept seperate, and if they run out of capacity due to a surge in COVID-19 admissions, additional capacity can be pooled from the elective and emergency capacity. <a href="https://pythonhealthdatascience.github.io/stars-reproduce-anagnostou-2022/evaluation/reproduction_success.html" target="_blank">Link to reproduction.</a>

**@johnson_cost_2021:** TBC <!-- TODO: Add Johnson -->

## Scope

| Study | Scope | Success | Time |
| --- | --- | --- | --- |
| Shoaib and Ramamohan 2022 | 17 items:<br>• 1 table<br>• 9 figures<br>• 7 in-text results | 16 out of 17 (94%) | 28h 14m |
| Huang et al. 2019 | 8 items:<br>• 5 figures<br>• 3 in-text results | 3 out of 8 (37.5%) | 24h 10m |
| Lim et al. 2020 | 9 items:<br>• 5 tables<br>• 4 figures | 9 out of 9 (100%) | 12h 27m |
| Kim et al. 2021 | 10 items:<br>• 3 tables<br>• 6 figures<br>• 1 in-text result | 10 out of 10 (100%) | 13h 59m |
| Anagnostou et al. 2022 | 1 item:<br>• 1 figure | 1 out of 1 (100%) | 2h 10m |
| Johnson et al. 2021 | 5 items:<br>• 1 table<br>• 4 figures | TBC | TBC |

## Time to completion

<!-- TODO: Add Johnson et al. 2021 once conducted reproduction -->

```{python}
# Import required libraries and functions
import matplotlib.pyplot as plt
import matplotlib.ticker as mtick
import numpy as np
import pandas as pd
from plotly import graph_objects as go
from functions import process_times, success_static, success_interactive

# Create a dataframe of times from each of the reproductions

# Shoaib and Ramamohan 2022 (16/17)
time_sho = [
    [856, 'Table 6'],
    [971, 'Figure 2A'],
    [889, 'Figure 2B'],
    [910, 'Figure 2C'],
    [981, 'Figure 2D'],
    [1002, 'Figure 3A'],
    [1008, 'Figure 3B'],
    [1009, 'Figure 3C'],
    [1012, 'Figure 3D'],
    [1165, 'Figure 4'],
    [1196, 'In-text result 1'],
    [1274, 'In-text result 2'],
    [1383, 'In-text result 3'],
    [1396, 'In-text result 4'],
    [1557, 'In-text result 5'],
    [1490, 'In-text result 6'],
    [1694, 'In-text result 7']]
df_sho = process_times(time_sho, adjust='In-text result 7')

# Huang et al. 2019 (3/8)
time_hua = [
    [516, 'In-text result 1'],
    [1193, 'In-text result 2'],
    [1450, 'Figure 2'], # Finished work, not completed, adjust to set to NaN
    [np.nan, 'Figure 3'],
    [np.nan, 'Figure 4'],
    [1228, 'Figure 5'],
    [np.nan, 'Supplementary figure'],
    [np.nan, 'In-text result 3']]
df_hua = process_times(time_hua, adjust='Figure 2')

# Lim et al. 2020 (9/9)
time_lim = [
    [607, 'Figure 2'],
    [647, 'Figure 3'],
    [655, 'Figure 4'],
    [747, 'Figure 5'],
    [486, 'Supplemental table 2'],
    [513, 'Supplemental table 3'],
    [513, 'Supplemental table 4'],
    [714, 'Supplemental table 5'],
    [690, 'Supplemental table 6']]
df_lim = process_times(time_lim)

# Kim et al. 2021 (10/10)
time_kim = [
    [445, 'Figure 1'],
    [511, 'Table 2'],
    [581, 'Figure 2'],
    [619, 'Figure 3'],
    [774, 'In-text result 1'],
    [672, 'Figure 4'],
    [785, 'Figure 5'],
    [793, 'Table 3'],
    [875, 'Supplementary figure 3'],
    [839, 'Supplementary table 2']]
df_kim = process_times(time_kim)

# Anagnostou et al. 2022 (1/1)
time_ana = [
    [130, 'Figure 3']]
df_ana = process_times(time_ana)

# Johnson et al. 2021
# TODO: Add times once completed reproduction

# Create list with the dataframes, labels and colours, to use for each plot
to_plot = [
    (df_sho, 'Shoaib and<br>Ramamohan<br>2022', '#ee6677'),
    (df_hua, 'Huang et<br>al. 2019', '#228833'),
    (df_lim, 'Lim et<br>al. 2020', '#4477aa'),
    (df_kim, 'Kim et<br>al. 2021', '#ccbb44'),
    (df_ana, 'Anagnostou<br>et al. 2022', '#66ccee')
    # 'Johnson et al. 2021', '#aa3377')
]
```

Non-interactive figure:

```{python}
# Create figure object
fig, ax = plt.subplots()

# Loop through and add lines for each reproduction
for x in to_plot:
    success_static(x[0], ax, x[1], x[2])

# Amend appearance
plt.xlim(0, 30)
plt.ylim(0, 100)
plt.xlabel('Time elapsed (hours)')
plt.ylabel('Percentage of items reproduced')
ax.yaxis.set_major_formatter(mtick.PercentFormatter())
ax.legend(loc='center left', bbox_to_anchor=(1, 0.5))

# Show figure
plt.show()
```

Interactive figure:

```{python}
# Create figure object
fig = go.Figure()

# Loop through and add lines for each reproduction
for x in to_plot:
    success_interactive(x[0], fig, x[1], x[2])

# Amend appearance
fig.update_layout(
    xaxis=dict(
        title='Time elapsed (hours)',
        range=[0, 30]
    ),
    yaxis=dict(
        title='Percentage of items reproduced',
        range=[0, 100],
        ticksuffix='%'
    ),
    showlegend=True)

# Show plot, hiding toolbar and not allowing zoom
fig.layout.xaxis.fixedrange = True
fig.layout.yaxis.fixedrange = True
fig.show(config={'displayModeBar': False})
```

## References