### 2024-09-10 pages/relationship_reproduction_evaluation.qmd
docs(relationship): add blank page to be completed with info on any relationship between reproduction and evaluation

---
title: "Relationship between reproduction and evaluation"
bibliography: ../references.bib
---

TODO: Add this page, to assess whether (and if so, how) there is a relationship between the evaluation (code and/or reporting) and how easy it was / able we were to reproduce