### 2024-10-25 pages/repo_evaluation.qmd
docs(repo evaluation): add johnson 2021

---
title: "Evaluation of the repository"
echo: False
bibliography: ../references.bib
---

The code and related research artefacts in the original code repositories were evaluated against:

* The criteria of badges related to reproducibility from various organisations and journals.
* Recommendations from the STARS framework for the sharing of code and associated materials from discrete-event simulation models (@monks_towards_2024).

Between each journal badge, there was often alot of overlap in criteria. Hence, a list of unique criteria was produced. The repositories are evaluated against this criteria, and then depending on which criteria they met, against the badges themselves.

*Caveat: Please note that these criteria are based on available information about each badge online, and that we have likely differences in our procedure (e.g. allowed troubleshooting for execution and reproduction, not under tight time pressure to complete). Moreover, we focus only on reproduction of the discrete-event simulation, and not on other aspects of the article. We cannot guarantee that the badges below would have been awarded in practice by these journals.*

Consider: **What criteria are people struggling to meet from the guidelines?**

<!-- TODO: Once complete, carefully go over again and check everything is correct -->

```{python}
# Imports and function
import plotly.express as px
import pandas as pd
from functions import eval_chart, plot_scatter
```

## Summary

**Unique badge criteria:**

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
criteria_dict = {
  'Shoaib and Ramamohan<br>2021 (16/17)': [3, 0, 9, 0],
  'Huang et al. 2019 (3/8)': [3, 0, 9, 0],
  'Lim et al. 2020 (9/9)': [3, 0, 9, 0],
  'Kim et al. 2021 (10/10)': [5, 0, 7, 0],
  'Anagnostou et al. 2022 (1/1)': [10, 0, 2, 0],
  'Johnson et al. 2021 (4/5)': [4, 0, 8, 0],
  'Hernandez et al. 2015 (1/8)': [3, 0, 9, 0],
  'Wood et al. 2021 (5/5)': [4, 0, 8, 0]
}
criteria_wide = pd.DataFrame(criteria_dict, index=col).T

eval_chart(criteria_wide)
```

::: {.callout icon=false collapse=false}

## Reflections

Four studies met 25% of criteria, and ranged from 12.5% to 100% reproduced.

Two studies met 33% of criteria, and these were 80% and 100% reproduced.

The remaining two studies were fully reproduced and met 41.7% and 83.3% of criteria.

However, I think it is more meaningful to actually look at what criteria were and were not met.

```{python}
plot_scatter('Unique badge criteria met',
             sho = 25,
             hua = 25,
             lim = 25,
             kim = 41.7,
             ana = 83.3,
             joh = 33.3,
             her = 25,
             woo = 33.3)
```

:::

**Badges:**

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
badge_dict = {
  'Shoaib and Ramamohan<br>2021 (16/17)': [3, 0, 9, 0],
  'Huang et al. 2019 (3/8)': [0, 0, 12, 0],
  'Lim et al. 2020 (9/9)': [3, 0, 9, 0],
  'Kim et al. 2021 (10/10)': [3, 0, 9, 0],
  'Anagnostou et al. 2022 (1/1)': [5, 0, 7, 0],
  'Johnson et al. 2021 (4/5)': [3, 0, 9, 0],
  'Hernandez et al. 2015 (1/8)': [0, 0, 12, 0],
  'Wood et al. 2021 (5/5)': [5, 0, 7, 0]
}
badge_wide = pd.DataFrame(badge_dict, index=col).T

eval_chart(badge_wide)
```

::: {.callout icon=false collapse=false}

## Reflections

Not certain how meaningful these numbers are, as we have imbalanced numbers of different types of badge, and meeting certain popular criteria will weight what is met v.s. not.

Feel that looking at the criteria met is a bit more meaningful? And then specific examples of how that translates into badges - e.g.

* None meeting ACM "Artifacts Evaluated - Functional" as requires xyz and these are commonly not met.
* For several, they meet three badges, but those three badges have one criteria: reproducing results.

Also, it's important to remember here that the criteria used for these were based on what could be found about each badge online, but we likely differences in our procedure (e.g. allowed troubleshooting for execution and reproduction, not under tight time pressure to complete). Moreover, we focus only on reproduction of the discrete-event simulation, and not on other aspects of the article. We cannot guarantee that the badges below would have been awarded in practice by these journals (and, in fact, know likely not for many).

```{python}
plot_scatter('Badges awarded',
             sho = 25,
             hua = 0,
             lim = 25,
             kim = 25,
             ana = 41.7,
             joh = 25,
             her = 0,
             woo = 41.7)
```

:::

**Essential components of STARS framework:**

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
essential_dict = {
  'Shoaib and Ramamohan<br>2021 (16/17)': [2, 0, 6, 0],
  'Huang et al. 2019 (3/8)': [2, 0, 6, 0],
  'Lim et al. 2020 (9/9)': [2, 0, 6, 0],
  'Kim et al. 2021 (10/10)': [4, 1, 3, 0],
  'Anagnostou et al. 2022 (1/1)': [7, 0, 1, 0],
  'Johnson et al. 2021 (4/5)': [4, 2, 2, 0],
  'Hernandez et al. 2015 (1/8)': [2, 0, 6, 0],
  'Wood et al. 2021 (5/5)': [2, 0, 6, 0]
}
essential_wide = pd.DataFrame(essential_dict, index=col).T

eval_chart(essential_wide)
```

::: {.callout icon=false collapse=false}

## Reflections

Similar to journal criteria (unsurprisingly, as looking at similar things) - most studies meet very few and have wide range of reproduction success, from 12.5% to 100%. Three met more, and these were 80% to 100% reproduced.

I think, if we were to draw anything from this, it would be to reflect on exactly what criteria were and were not met, and why/how that impacted reproduction, in any way (either success or time).

*Note: Just considers those fully met, in plot*

```{python}
plot_scatter('Essential STARS components met',
             sho = 25,
             hua = 25,
             lim = 25,
             kim = 50,
             ana = 87.5,
             joh = 50,
             her = 25,
             woo = 25)
```

:::

**Optional components of STARS framework:**

```{python}
# Create dataframe of results
optional_dict = {
  'Shoaib and Ramamohan<br>2021 (16/17)': [0, 0, 5, 0],
  'Huang et al. 2019 (3/8)': [2, 0, 3, 0],
  'Lim et al. 2020 (9/9)': [0, 0, 5, 0],
  'Kim et al. 2021 (10/10)': [0, 0, 5, 0],
  'Anagnostou et al. 2022 (1/1)': [1, 1, 3, 0],
  'Johnson et al. 2021 (4/5)': [0, 0, 5, 0],
  'Hernandez et al. 2015 (1/8)': [0, 0, 5, 0],
  'Wood et al. 2021 (5/5)': [0, 0, 5, 0]
}
optional_wide = pd.DataFrame(optional_dict, index=col).T

eval_chart(optional_wide)
```

::: {.callout icon=false collapse=false}

## Reflections

This highlights how Huang meets the most criteria, but is only partially reproduced - but I think it is most interesting to consider why this is.

*Note: Just considers those fully met, in plot*

```{python}
plot_scatter('Optional STARS components met',
             sho = 0,
             hua = 40,
             lim = 0,
             kim = 0,
             ana = 20,
             joh = 0,
             her = 0,
             woo = 0)
```

:::

## Journal badges

Key:

* **S:** @shoaib_simulation_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-shoaib-2022/evaluation/badges.html" target="_blank">link to evaluation</a>
* **Hu:** @huang_optimizing_2019  - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-huang-2019/evaluation/badges.html" target="_blank">link to evaluation</a>
* **L:** @lim_staff_2020 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-lim-2020/evaluation/badges.html" target="_blank">link to evaluation</a>
* **K:** @kim_modelling_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-kim-2021/evaluation/badges.html" target="_blank">link to evaluation</a>
* **A:** @anagnostou_facs-charm_2022 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-anagnostou-2022/evaluation/badges.html" target="_blank">link to evaluation</a>
* **J:** @johnson_cost_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-johnson-2021/evaluation/badges.html" target="_blank">link to evaluation</a>
* **He:** @hernandez_optimal_2015 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-hernandez-2015/evaluation/badges.html" target="_blank">link to evaluation</a>
* **W:** @wood_value_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-wood-2021/evaluation/badges.html" target="_blank">link to evaluation</a>

<!-- TODO: Add Johnson et al. 2021 -->

In this section and below, the criteria for each study are marked as either being fully met (✅), partially met (🟡), not met (❌) or not applicable (N/A).

**Unique criteria:**

| Item | S | Hu | L | K | A | J | He | W |
| - | - | - | - | - | - | - | - | - | - |
| **Criteria related to how artefacts are shared** |
| Stored in a permanent archive that is publicly and openly accessible | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| Has a persistent identifier | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| Includes an open licence | ❌ | ✅ | ❌ | ✅ | ✅ | ✅ | ❌ | ❌ |
| **Criteria related to what artefacts are shared** |
| Artefacts are relevant to and contribute to the article’s results | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ |
| Complete set of materials shared (as would be needed to fully reproduce article) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ✅ |
| **Criteria related to the structure and documentation of the artefacts** |
| Artefacts are well structured/organised (e.g. to the extent that reuse and repurposing is facilitated, adhering to norms and standards of research community) | ❌ | ❌ | ❌ | ✅ | ✅ | ❌ | ✅ | ❌ |
| Artefacts are sufficiently documented (i.e. to understand how it works, to enable it to be run, including package versions) | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| Artefacts are carefully documented (more than sufficient - i.e. to the extent that reuse and repurposing is facilitated - e.g. changing parameters, reusing for own purpose) | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| Artefacts are clearly documented and accompanied by a README file with step-by-step instructions on how to reproduce results in the manuscript | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| **Criteria related to running and reproducing results** |
| Scripts can be successfully executed | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ |
| Independent party regenerated results using the authors research artefacts | ✅ | ❌ | ✅ | ✅ | ✅ | ✅ | ❌ | ✅ |
| Reproduced within approximately one hour (excluding compute time) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
: {tbl-colwidths="[60, 5, 5, 5, 5, 5, 5, 5, 5]"}

::: {.callout icon=false collapse=false}

## Reflections

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
criteria_dict = {
  'Archive': [1, 0, 7, 0],
  'ID': [1, 0, 7, 0],
  'Licence': [4, 0, 4, 0],
  'Relevant': [8, 0, 0, 0],
  'Complete': [1, 0, 7, 0],
  'Structured for reuse': [3, 0, 5, 0],
  'Sufficient docs': [1, 0, 7, 0],
  'Careful docs': [1, 0, 7, 0],
  'Step-by-step README': [1, 0, 7, 0],
  'Execute': [8, 0, 0, 0],
  'Reproduced': [6, 0, 2, 0],
  'One hour': [0, 0, 8, 0]
}
criteria_wide = pd.DataFrame(criteria_dict, index=col).T

eval_chart(criteria_wide)
```

**Stored in a permanent archive that is publicly and openly accessible**: Fulfillment doesn't impact reproduction as I was able to get everything needed from the remote code repository (GitHub or GitLab). However, if these had been deleted from GitHub, it would have become invaluable.

**Has a persistent identifier**: Fulfillment doesn't impact reproduction.

**Includes an open licence**: This had a big impact on our ability to complete reproductions, as we had to ask authors to add an open licence to their work, to enable us to use it. Gladly, all authors we contacted kindly add these on request. However, it's worth noting that this was a relatively common issue, and one of the most important, since it completely prevents reuse if excluded.

**Artefacts are relevant to and contribute to the article’s results**: All met (if not met, this would be a massive hindrance).

**Complete set of materials shared (as would be needed to fully reproduce article)**: This had a really big impact on the reproduction. The main reason for longer times in reproduction was (a) code for scenarios not provided, and (b) code to process results into figures and tables not provided.

**Artefacts are well structured/organised (e.g. to the extent that reuse and repurposing is facilitated, adhering to norms and standards of research community)**: In cases where this was not met, this was often related to hard-coding of parameters, or set up of code in a way that - alike hard-coding - made it much harder to reuse the model (in this context, to reuse them between different scenarios) - or quite a busy/cluttered repository which was confusing to navigate.

**Artefacts are sufficiently documented (i.e. to understand how it works, to enable it to be run, including package versions)**: Only three studies had any documentation - READMEs for @kim_modelling_2021, @anagnostou_facs-charm_2022 and @johnson_cost_2021 - but this criteria also required "package versions" which were uncommon. However, focusing on the READMEs, in each of these cases it was great to have these, guiding on how to run the scripts, or on what each folder/file in the repository is.

**Artefacts are carefully documented (more than sufficient - i.e. to the extent that reuse and repurposing is facilitated - e.g. changing parameters, reusing for own purpose)**: In @anagnostou_facs-charm_2022, they include a file `CHARM_INFO.md` alongside their README which walks through the input parameters for the model. I didn't need to change any of these for the reproduction, but would imagine this is to be very helpful if someone were to reuse the model.

**Artefacts are clearly documented and accompanied by a README file with step-by-step instructions on how to reproduce results in the manuscript**: Whilst this was true for @anagnostou_facs-charm_2022, it should be noted that this was a very simple example, just requiring to run one script which quickly reproduces everything! I had been a bit uncertain on it, since the README doesn’t explicitly say how to make the figure, but it does provide instructions that lead you to regenerate the exact model results from the paper, and so I feel that it does provide instructions to reproduce results sufficiently (although would be more complete to include instructions for figure too - so if it weren’t a yes/no decision for badges, I would’ve said this was partially met). Ideally, studies would clearly outline how to reproduce results in full.

**Scripts can be successfully executed**: This is true, though I did allow troubleshooting, which sometimes took a long while. Hence, the importance of e.g. environments and scripts being provided in a runnable format (both covered on the [reflections page](reflections.qmd)), since these are the hurdles to successfully executing scripts.

**Independent party regenerated results using the authors research artefacts**: On the [reproduction page](reproduction.qmd), I reflected (where possible) on what I thought the primary reasons were, for cases where I didn't manage to reproduce results despite troubleshooting.

**Reproduced within approximately one hour (excluding compute time)**: In this study, this was pretty much impossible, since I followed a protocol of first setting up, reading the article, and so on. It is worth noting however that there were two studies that were quite quick to run, which I reflect about on the [reproduction page](reproduction.qmd).

:::

**Badges:**

The badges are grouped into three categories:

* "Open objects" badges: These badges relate to research artefacts being made openly available.
* "Object review" badges: These badges relate to the research artefacts being reviewed against criteria of the badge issuer.
* "Reproduced" badges: These badges relate to an independent party regenerating the reuslts of the article using the author objects.

<!-- TODO: Add Johnson et al. 2021 -->

| Item | S | Hu | L | K | A | J | He | W |
| - | - | - | - | - | - | - | - | - | - |
| **“Open objects” badges** |
| **NISO "Open Research Objects (ORO)"**<br>• Stored in a permanent archive that is publicly and openly accessible<br>• Has a persistent identifier<br>• Includes an open licence | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| **NISO "Open Research Objects - All (ORO-A)"**<br>• Stored in a permanent archive that is publicly and openly accessible<br>• Has a persistent identifier<br>• Includes an open licence<br>• Complete set of materials shared (as would be needed to fully reproduce article) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **ACM "Artifacts Available"**<br>• Stored in a permanent archive that is publicly and openly accessible<br>• Has a persistent identifier | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| **COS "Open Code"**<br>• Stored in a permanent archive that is publicly and openly accessible<br>• Has a persistent identifier<br>• Includes an open licence<br>• Complete set of materials shared (as would be needed to fully reproduce article)<br>• Artefacts are sufficiently documented (i.e. to understand how it works, to enable it to be run, including package versions) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **IEEE "Code Available"**<br>• Complete set of materials shared (as would be needed to fully reproduce article) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ✅ |
| **"Object review" badges** |
| **ACM "Artifacts Evaluated - Functional"**<br>• Artefacts are sufficiently documented (i.e. to understand how it works, to enable it to be run, including package versions)<br>• Artefacts are relevant to and contribute to the article’s results<br>• Complete set of materials shared (as would be needed to fully reproduce article)<br>• Scripts can be successfully executed | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **ACM "Artifacts Evaluated - Reusable"**<br>• Artefacts are sufficiently documented (i.e. to understand how it works, to enable it to be run, including package versions)<br>• Artefacts are carefully documented (more than sufficient - i.e. to the extent that reuse and repurposing is facilitated - e.g. changing parameters, reusing for own purpose)<br>• Artefacts are relevant to and contribute to the article's results<br>• Complete set of materials shared (as would be needed to fully reproduce article)<br>• Scripts can be successfully executed<br>• Artefacts are well structured/organised (e.g. to the extent that reuse and repurposing is facilitated, adhering to norms and standards of research community) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **IEEE "Code Reviewed"**<br>• Complete set of materials shared (as would be needed to fully reproduce article)<br>• Scripts can be successfully executed | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ✅ |
| **"Reproduced" badges** |
| **NISO "Results Reproduced (ROR-R)"**<br>• Independent party regenerated results using the authors research artefacts | ✅ | ❌ | ✅ | ✅ | ✅ | ✅ | ❌ | ✅ |
| **ACM "Results Reproduced"**<br>• Independent party regenerated results using the authors research artefacts | ✅ | ❌ | ✅ | ✅ | ✅ | ✅ | ❌ | ✅ |
| **IEEE "Code Reproducible"**<br>• Independent party regenerated results using the authors research artefacts | ✅ | ❌ | ✅ | ✅ | ✅ | ✅ | ❌ | ✅ |
| **Psychological Science "Computational Reproducibility"**<br>• Independent party regenerated results using the authors research artefacts<br>• Reproduced within approximately one hour (excluding compute time)<br>• Artefacts are well structured/organised (e.g. to the extent that reuse and repurposing is facilitated, adhering to norms and standards of research community)<br>• Artefacts are clearly documented and accompanied by a README file with step-by-step instructions on how to reproduce results in the manuscript | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
: {tbl-colwidths="[60, 5, 5, 5, 5, 5, 5, 5, 5]"}

::: {.callout icon=false collapse=false}

## Reflections

<!-- TODO: Update after Johnson et al. 2021 -->

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
badges_dict = {
  'NISO “Open Research<br>Objects (ORO)”': [1, 0, 7, 0],
  'NISO “Open Research<br>Objects - All (ORO-A)”': [0, 0, 8, 0],
  'ACM “Artifacts<br>Available”': [1, 0, 7, 0],
  'COS “Open Code”': [0, 0, 8, 0],
  'IEEE “Code Available”': [1, 0, 7, 0],
  'ACM “Artifacts<br>Evaluated - Functional”': [0, 0, 8, 0],
  'ACM “Artifacts<br>Evaluated - Reusable”': [0, 0, 8, 0],
  'IEEE “Code Reviewed”': [1, 0, 7, 0],
  'NISO “Results<br>Reproduced (ROR-R)”': [6, 0, 2, 0],
  'ACM “Results Reproduced”': [6, 0, 2, 0],
  'IEEE “Code Reproducible”': [6, 0, 2, 0],
  'Psychological Science<br>“Computational Reproducibility”': [0, 0, 8, 0],
}
badges_wide = pd.DataFrame(badges_dict, index=col).T

eval_chart(badges_wide)
```

Only one study had **permanent archive (with persistent identifier)**, hence one being awarded *NISO “Open Research Objects (ORO)”* and *ACM “Artifacts Available”*. This was required by two other badges **as well as** sufficient documentation and/or complete set of materials. Since the one permanently archived study didn't meet these criteria, none were awarded the two badges: *NISO “Open Research Objects - All (ORO-A)”* or *COS “Open Code”*.

A **complete set of materials** was required by *IEEE “Code Available”* and *IEEE “Code Reviewed”* - but this was only met by one study, as studies commonly did not include code for scenarios or creation of figures and tables. It was also required by *ACM “Artifacts Evaluated - Functional and Reusable”* badges, but since that one study didn't meet their **documentation** requirements, none were awarded those badges.

Three badges had one criteria: **reproduction of results** - and hence, several studies received these (*NISO “Results Reproduced (ROR-R)”*, *ACM “Results Reproduced”*, *IEEE “Code Reproducible”*). However, it's worth noting that we allowed troubleshooting (since that was how we approached the reproductions), and that some of these studies might not have received these badges, if they have any requirements that exclude troubleshooting (which is likely).

The final badge *Psychological Science “Computational Reproducibility”* had several criteria, but of importance, one of those was to complete the **reproduction within an hour**. This was somewhat impossible in our procedures, since we read the article and set-up etc beforehand, and so none were awarded this badge.

:::

## STARS framework

Key:

* **S:** @shoaib_simulation_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-shoaib-2022/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **Hu:** @huang_optimizing_2019  - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-huang-2019/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **L:** @lim_staff_2020 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-lim-2020/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **K:** @kim_modelling_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-kim-2021/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **A:** @anagnostou_facs-charm_2022 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-anagnostou-2022/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **J:** @johnson_cost_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-johnson-2021/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **He:** @hernandez_optimal_2015 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-hernandez-2015/evaluation/artefacts.html" target="_blank">link to evaluation</a>
* **W:** @wood_value_2021 - <a href="https://pythonhealthdatascience.github.io/stars-reproduce-wood-2021/evaluation/artefacts.html" target="_blank">link to evaluation</a>

<!-- TODO: Add Johnson et al. 2021 -->

| Item | S | Hu | L | K | A | J | He | W |
| - | - | - | - | - | - | - | - | - | - |
| **Essential components** |
| **Open licence**<br>Free and open-source software (FOSS) licence (e.g. MIT, GNU Public Licence (GPL)) | ❌ | ✅ | ❌ | ✅ | ✅ | ✅ | ❌ | ❌ |
| **Dependency management**<br>Specify software libraries, version numbers and sources (e.g. dependency management tools like virtualenv, conda, poetry) | ❌ | ❌ | ❌ | 🟡 | ✅ | 🟡 | ❌ | ❌ |
| **FOSS model**<br>Coded in FOSS language (e.g. R, Julia, Python) | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ |
| **Minimum documentation**<br>Minimal instructions (e.g. in README) that overview (a) what model does, (b) how to install and run model to obtain results, and (c) how to vary parameters to run new experiments | ❌ | ❌ | ❌ | ✅ | ✅ | 🟡 | ❌ | ❌ |
| **ORCID**<br>ORCID for each study author | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **Citation information**<br>Instructions on how to cite the research artefact (e.g. CITATION.cff file) | ❌ | ❌ | ❌ | ❌ | ✅ | ✅ | ❌ | ❌ |
| **Remote code repository**<br>Code available in a remote code repository (e.g. GitHub, GitLab, BitBucket) | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ | ✅ |
| **Open science archive**<br>Code stored in an open science archive with FORCE11 compliant citation and guaranteed persistance of digital artefacts (e.g. Figshare, Zenodo, the Open Science Framework (OSF), and the Computational Modeling in the Social and Ecological Sciences Network (CoMSES Net)) | ❌ | ❌ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| **Optional components** |
| **Enhanced documentation**<br>Open and high quality documentation on how the model is implemented and works  (e.g. via notebooks and markdown files, brought together using software like Quarto and Jupyter Book). Suggested content includes:<br>• Plain english summary of project and model<br>• Clarifying licence<br>• Citation instructions<br>• Contribution instructions<br>• Model installation instructions<br>• Structured code walk through of model<br>• Documentation of modelling cycle using TRACE<br>• Annotated simulation reporting guidelines<br>• Clear description of model validation including its intended purpose | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **Documentation hosting**<br>Host documentation (e.g. with GitHub pages, GitLab pages, BitBucket Cloud, Quarto Pub) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **Online coding environment**<br>Provide an online environment where users can run and change code (e.g. BinderHub, Google Colaboratory, Deepnote) | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ | ❌ |
| **Model interface**<br>Provide web application interface to the model so it is accessible to less technical simulation users | ❌ | ✅ | ❌ | ❌ | ✅ | ❌ | ❌ | ❌ |
| **Web app hosting**<br>Host web app online (e.g. Streamlit Community Cloud, ShinyApps hosting) | ❌ | ✅ | ❌ | ❌ | 🟡 | ❌ | ❌ | ❌ |
: {tbl-colwidths="[60, 5, 5, 5, 5, 5, 5, 5, 5]"}

::: {.callout icon=false collapse=false}

## Reflections

<!-- TODO: Update after Johnson et al. 2021 -->

```{python}
# Create dataframe of results
col = ['fully', 'partially', 'not', 'na']
stars_dict = {
  'Open licence': [4, 0, 4, 0],
  'Dependency management': [1, 2, 5, 0],
  'FOSS model': [8, 0, 0, 0],
  'Minimum documentation': [2, 1, 5, 0],
  'ORCID': [0, 0, 8, 0],
  'Citation information': [2, 0, 6, 0],
  'Remote code repository': [8, 0, 0, 0],
  'Open science archive': [1, 0, 7, 0],
  'Enhanced documentation': [0, 0, 8, 0],
  'Documentation hosting': [0, 0, 8, 0],
  'Online coding environment': [0, 0, 8, 0],
  'Model interface': [2, 0, 6, 0],
  'Web app hosting': [1, 1, 6, 0],
}
stars_wide = pd.DataFrame(stars_dict, index=col).T

eval_chart(stars_wide)
```

These topics were covered in the badge criteria reflections: **open licence**, **minimum documentation**, and **open science archive**. <!--TODO: Consider re minimum documentation: Kim met for STARS but not for badges. Double-check that decision-->

**Dependency management**: This was pretty uncommon, and often took some troubleshooting at the start, to figure out which packages were needed, and certain versions.

**FOSS model**: All met as requirement of our reproduction.

**ORCID** and **citation information**: Doesn't impact reproduction in this case - but we do go to these from having found an article. In all cases, I emailed the authors, which requires finding contact information (generally via paper, sometimes from googling them to find new emails).

**Remote code repository**: All met, most common way to share code.

**Enhanced documentation**: Only three studies had any documentation, and neither met these extensive requirements. I anticipate - if any had met this - it would've made the reproduction very quick and easy!

**Documentation hosting**: Not applicable, given only basic documentation.

**Online coding environment**: None provided. I always intended to run on my own machine, so this might not have had much bearing in my case if provided, but would moreso for people who perhaps didn't have Python or R installed, and hopefully would have bypassed environment troubleshooting issues.

**Model interface**: Two studies had applications, although in both cases, these weren't "outcomes" in scope of reproduction, nor did they produce them.

**Web app hosting**: This was quite important. Both apps had been hosted, but one was hosted with a site that is no longer operational. In both cases, the app wasn't in "scope" although I did still view it and look into it for one as it was hosted and so could very easily - but for the other, I didn't view it, as I didn't go through the steps of running it locally, since it wasn't the focus.

:::

## Timings

* @shoaib_simulation_2021 - 30m
* @huang_optimizing_2019 - 17m
* @lim_staff_2020 - 18m
* @kim_modelling_2021 - 18m
* @anagnostou_facs-charm_2022 - 19m
* @johnson_cost_2021 - 20m
* @hernandez_optimal_2015 - 13m
* @wood_value_2021 - 14m

::: {.callout icon=false collapse=false}

## Reflections

No particular comments, don't think we learn much from the timings here.

:::

## Badge sources

**National Information Standards Organisation (NISO)** (@niso_reproducibility_badging_and_definitions_working_group_reproducibility_2021)

* "Open Research Objects (ORO)"
* "Open Research Objects - All (ORO-A)"
* "Results Reproduced (ROR-R)"

**Association for Computing Machinery (ACM)** (@association_for_computing_machinery_acm_artifact_2020)

* "Artifacts Available"
* "Artifacts Evaluated - Functional"
* "Artifacts Evaluated - Resuable"
* "Results Reproduced"

**Center for Open Science (COS)** (@blohowiak_badges_2023)

* "Open Code"

**Institute of Electrical and Electronics Engineers (IEEE)** (@institute_of_electrical_and_electronics_engineers_ieee_about_nodate)

* "Code Available"
* "Code Reviewed"
* "Code Reproducible"

**Psychological Science** (@hardwicke_transparency_2023 and @association_for_psychological_science_aps_psychological_2023)

* "Computational Reproducibility"

## References