### 2024-10-25 index.qmd
chore(docs): correct licence spelling

---
title: Overview
authors:
  - name: Ms Amy Heather
    url: https://github.com/amyheather
    orcid: 0000-0002-6596-3479
    affiliation: Peninsula Collaboration for Health Operational Research and Data Science (PenCHORD), University of Exeter Medical School
  - name: Associate Professor Tom Monks
    url: https://github.com/TomMonks
    orcid: 0000-0003-2631-4481
    affiliation: Peninsula Collaboration for Health Operational Research and Data Science (PenCHORD), University of Exeter Medical School
  - name: Dr Alison Harper
    url: https://github.com/AliHarp
    orcid: 0000-0001-5274-5037
    affiliation: Centre for Simulation, Analytics and Modelling (CSAM), University of Exeter Business School
---

## Overview {.unnumbered}

This book describes the findings from work package 1 of the project **STARS: Sharing Tools and Artefacts for Reproducible Simulations in healthcare**.

Use the sidebar to navigate through:

* **Introduction** - background on STARS, pilot work, and the aim of work package 1
* **Methods** - summarises the methods which are described in detail in our protocol
* **Results** - describes the results of the reproductions and evaluations, and reflections from the process
* **Discussion** - considers and applies findings \

## Funding {.unnumbered}

This work is supported by the Medical Research Council [grant number [MR/Z503915/1](https://gtr.ukri.org/projects?ref=MR%2FZ503915%2F1)].

## Licence {.unnumbered}

This book is licensed under a [Creative Commons Attribution 4.0 International (CC BY 4.0)](https://github.com/pythonhealthdatascience/stars_wp1_summary/blob/1a7e932013abd8e4810764b8925e98e9a6473844/LICENSE) licence.

[![Licence: CC BY 4.0](https://licensebuttons.net/l/by/4.0/80x15.png)](https://creativecommons.org/licenses/by/4.0/)
[![Licence: CC BY 4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)